---
title: "p8105_hw6_nz2333"
author: "nz2333"
date: "12/4/2021"
output: github_document
---

```{r}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

_Clean data_
```{r}
problem1_df = read_csv("./birthweight.csv") %>%
  janitor::clean_names() %>%
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace),
    
    babysex = recode(babysex, "1" = "Male", "2" = "Female"), 
    frace = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown"),
    malform = recode(malform, "0" = "absent", "1" = "present"),
    mrace = recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown")
  )

problem1_df
```

_Check for missing values_
```{r}
map(problem1_df, ~sum(is.na(.)))
```
No missing data is found for all variables. 

_Model proposal_

Since we are trying to understand the effects of several variables on a child’s birthweight, I propose to associate birthweight with gender, baby's length and head circumstances at birth, mother's weight gain during pregnancy and mother's height. 
```{r}
p1_fit = lm(bwt ~ babysex + bhead + blength + wtgain + mheight, data = problem1_df)
```

_Plot of residuals and predicted values_
```{r}
problem1_df %>% 
  modelr::add_residuals(p1_fit) %>%
  modelr::add_predictions(p1_fit) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.3) +
  geom_smooth(color = "red", method = "lm", se = FALSE) +
  labs(
    title = "Residual plot for predicted values", 
    x = "Predicted values", 
    y = "Residuals"
  )
```

Residuals are evenly distributed around 0, so the asusmption of constant variance of residuals hold. However, there are some extreme values on the left side, suggesting outliers. 

_Comparison with two other models_
```{r}
model1 = lm(bwt ~ blength + gaweeks, data = problem1_df)
model2 = lm(bwt ~ bhead + blength + babysex + bhead*blength + blength*babysex + bhead*babysex + bhead*blength*babysex, data = problem1_df)

broom::tidy(model1)
broom::tidy(model2)
```

_Cross validation_
```{r}
cv_df = crossv_mc(problem1_df, 100)

cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_df = cv_df %>%
  mutate(
    p1_fit = map(train, ~lm(bwt ~ babysex + bhead + blength + wtgain + mheight, data = .x)), 
    model1 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)), 
    model2 = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + blength*babysex + bhead*babysex + bhead*blength*babysex, data = .x))
  ) %>%
  mutate(
    rmse_p1fit = map2_dbl(p1_fit, test, ~rmse(model = .x, data = .y)),
    rmse_model1 = map2_dbl(model1, test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(model2, test, ~rmse(model = .x, data = .y)))
```

_Plot of RMSEs_
```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse, fill = model)) + 
  geom_violin(alpha = 0.3) +
  geom_boxplot(alpha = 0.4) +
  labs(
    title = "Violin plot and box plot for RMSE across three models", 
    x = "model", 
    y = "RMSE"
  )
```

The plot shows that my proposed model p1fit has the lowest RMSE, which means that this model is the best model fitted among the three through comparison. 
   
## Problem 2
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

_R squared_
```{r}
p2_r2 = weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results)

p2_r2 %>%
  summarize(
    ci_lower = quantile(r.squared, 0.025), 
    ci_upper = quantile(r.squared, 0.975))

p2_r2 %>%
ggplot(aes(x = r.squared)) +
  geom_density() +
  labs(
    title = "Distribution plot of R squared estimates", 
    x = "R-squared", 
    y = "Density"
  )
```

The 95% Confidence interval for R squared is (0.894, 0.927). 
The distribution over the R squared estimates approximates a normal distribution with a bell shaped curve. The mean is `r round(mean(pull(p2_r2, r.squared)), 3)` and standard deviation is `r round(sd(pull(p2_r2, r.squared)), 3)`. 

_Log betas_
```{r}
p2_log = weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results)

log = p2_log %>%
  select(.id, term, estimate) %>%
  pivot_wider(
    names_from = "term",
    values_from = "estimate"
  ) %>%
  mutate(
    log = log(`(Intercept)` *tmin)
  )

log %>%
  summarize(
    ci_lower = quantile(log, 0.025), 
    ci_upper = quantile(log, 0.975))

log %>%
ggplot(aes(x = log)) +
  geom_density() +
  labs(
    title = "Distribution plot of Log(beta0*beta1) estimates", 
    x = "Log(beta0*beta1)", 
    y = "Density"
  )
```


The 95% Confidence interval for Log(beta0*beta1) is (1.97, 2.06). 
The distribution over the Log(beta0*beta1) estimates approximates a normal distribution with a bell shaped curve. The mean is `r round(mean(pull(log, log)), 3)` and standard deviation is `r round(sd(pull(log, log)), 3)` 
